\bf A Comparision of Bootstrap Confidence Interval Methods

### 2022-12-08 project/PaperTemplate/paper.tex
Finished running simulations and edited once

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge
Profossor Wagaman} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

@@

\bigskip
\begin{abstract}
For my project, I plan to explore different bootstrap methods for
creating confidence intervals and then perform comparisons between a
couple of the methods. My paper will re-introduce the idea of bootstrap
to my peers and give some background on constructing confidence
intervals. We will also go deeper into the theory behind the
construction of confidence intervals from bootstrapped data. These
various methods to create a confidence interval from a bootstrap can
include the percentile method, bias-corrected method, accelerated
method, and the studentized method, as well as others. I will
demonstrate how these bootstrap methods work using ``toy examples,''
which will be datasets in which a specific bootstrap method is
appropriate. To demonstrate my understanding of the methods, I will
write a simulation to compare a few of them and determine how they
perform against each other. I will show my understanding of the
different methods of creating confidence intervals from bootstrapped
data by communicating the statistical theory in a concise and accessible
way to my peers. Writing the simulation and sharing conclusions will
demonstrate my ability to implement statistical methods in practice as
well as my ability to analyze the results of the simulation.
\end{abstract}

@@

Bootstrapping is an essential tool in Statistics, even more so now that
there is access to higher computing power. We use bootstrapping to
estimate the desired population parameter from a given sample without
making assumptions about any underlying distributions of the sample.
Different bootstrap techniques are developed, but we will focus on the
non-parametric bootstrap for our purposes. One way to make a statistical
inference is through confidence intervals which give a range of
estimates for the unknown population parameter at a certain confidence
level. We can use bootstrapping to create accurate approximate
confidence intervals for our population parameter even without its
underlying distributions. There are many different ways to create
intervals, and we will go through a couple of them: The Standard Method,
The Percentile Method, The Bias-Corrected (BC) Method, The
Bias-Corrected with Acceleration (BCa) Method, and finally, The
Studentized Method. Additionally, we will go through an example of
creating a bootstrapped confidence interval for each method. Then, we
will perform a simulation to evaluate the performance of the
bootstrapped confidence interval methods on different distribution
types.

@@

First, we introduce the non-parametric bootstrap. Suppose we have a
random sample \(X = (x_1,x_2,\dots,x_n)\) from our unknown distribution,
\(F\) and a statistic of interest \(\hat{\theta} = \hat{\theta}(X)\)
\citep{EfronCasi}. Ideally, the desired test statistic could be found by
repeatedly sampling new reproductions of \(X\) from \(F\). However \(F\)
is unknown, so this is not possible. The non-parametric bootstrap
creates an estimate \(\hat{F}\) from \(F\) using our sample \(X\)
without making any parametric assumptions about \(F\) (such as its
distribution type). Therefore, the bootstrap sample could be represented
as \(X^* = (x^*_1, x^*_2, \dots, x^*_n)\) where each \(x^*_i\) is
sampled randomly with equal probability and with replacement from
\(\{x_1,x_2,\dots,x_n\}\). From this bootstrap sample, a bootstrap
replication of the test statistic can be computed using
\(\hat{\theta}^* = \hat{\theta}(X^*)\). A large number, \(B\), of
bootstrap samples are drawn independently and the corresponding
bootstrap replication of the test statistic is calculated.
\[\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B.\]
The bootstrap estimate of the test statistic is the empirical value of
the test statistic from all of the \(\hat{\theta}^{*b}\) replications.
As \(B\) increases, \(\hat{F}\) approaches \(F\) which means that the
test statistic of interest approaches its true value as well.

We demonstrate performing a non-parametric bootstrap below using
\texttt{SnowGR} which gives the official snowfall dataset by month for
Grand Rapids, MI, starting in 1893. We will be using the \texttt{Dec}
variable which gives the number of inches of snow that fell in December
of each year. In particular, we are interested in the sample mean and
later finding confidence intervals for population mean of \texttt{Dec}.
The distribution is below.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{data}\NormalTok{(SnowGR)}
\FunctionTok{favstats}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  min   Q1 median    Q3  max     mean       sd   n missing
##  0.8 8.35   13.8 18.65 59.2 15.75798 10.58382 119       0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 1: Histogram with Studentized Interval Shown"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Next, we will perform the non-parametric bootstrap using 10,000
replications.

@@

Below is the

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_dhistogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\AttributeTok{data =}\NormalTok{ dec\_means) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{gf\_dens}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{gf\_labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Histogram of 10,000 Bootstrapped Mean Dec Values"}\NormalTok{,}
          \AttributeTok{caption =} \StringTok{"Fig. 2: Bootstrap Histogram"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. As such, the
standard confidence interval is sometimes called the normal confidence
interval for bootstrapped data. This means that for an unknown
distribution, the standard confidence interval could present an
incorrect range. However, the same process can be used with bootstrap
sampling to form the bootstrap percentile method. This means that an
approximate bootstrap confidence interval will be created in the same
automatic way that the standard confidence interval was created. For
bootstrapped confidence intervals, the number of bootstrap replications
\(B\) must be large (around 2000) due to the nature of confidence
intervals requiring greater accuracy \citep{Efron86}.

Using our example, to find a 95\% confidence interval for the population
mean of \texttt{Dec}, we would get the following:

@@

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}

@@

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\([\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}]\)
\citep[\citet{EfronCasi}]{Efron86}. To go further, we can define
\(\hat{G}(t)\) as the bootstrap cdf, or the proportion of bootstrap
samples less than \(t\):
\[\hat{G}(t) = \frac{\#\{\hat{\theta^{*b} \leq t}\}}{B}.\] Thus the
\(\alpha\)th percentile point of the distribution is given by
\[\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).\]
It follows that the percentile interval can be represented as
\[\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].\] In the
case that the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is not usually the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] This
also can be represented as an interval,
\[\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ].\]
Therefore, the percentile method produces a correct interval for
\(\phi\) and due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist \citep{Efron86}.

@@

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf} This
interval is similar to the one created using the Standard method, but
shifted slightly to the right.

@@

The next method we will be looking at is the bias-corrected percentile
method (BC method) which is an improvement upon the previous percentile
method because we now take into account the possibility of bias. It can
be shown that \(\hat{\theta}\) is biased upwards relative to \(\theta\)
which means that the confidence intervals should be adjusted downwards
\citep[\citet{EfronCasi}]{Efron86}. From our simulated bootstrap
replications
\(\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B},\)
define \[p_0 = \frac{\#\{\hat{\theta^{*b} \leq \theta}\}}{B},\] and
define the bias-correction value \[z_0 = \Phi^{-1}(p_0),\] where
\(\Phi^{-1}\) is the inverse function of the standard normal cdf. Thus
we define a transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\), \[\hat{\phi} \sim N(\phi - z_0\tau, \tau^2),\] with \(z_0\)
and \(\tau\) constants. This means that we can say the bias corrected
method has an \(\alpha\)-level endpoint can be represented as
\[\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].\]
If \(\hat{G} = 0.50\), then half of the bootstrap distribution is less
than \(\hat{\theta}\) and our bias-correction value \(z_0 = 0\). In this
case, the confidence interval produced by BC would be the same interval
produced by the percentile method.

Using this method to create a 95\% confidence interval has a couple more
steps because we need to find the bias-correction value. First we
calculate the sample mean.

@@

Then we can find the modified percentiles:

@@

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

A further modification upon the BC interval is the bias corrected and
accelerated method (BCa). For this method, we do not assume the the
standard error, \(\tau\) is constant as we did in the BC interval
\citep[\citet{EfronCasi}]{Efron86}. Rather, we assume the existence of a
monotone transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\),
\[\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2) \text{ where } \tau_\phi = 1+ a\phi.\]
The \(a\) is known as the acceleration and is a constant that describes
how the standard deviation of \(\hat{\phi}\) varies with \(\phi\). In
other words, \(a\) is proportional to the skewness of the bootstrap
distribution. For example, for one-parameter exponential families,
\(a=z_0\), however, there are many different algorithms to compute and
estimate \(a\) \citep{Flowers18}. Now, our \(\alpha\)-level endpoint
from BCa is
\[\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].\]
If \(a = 0\), then
\(\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].\) When
calculating a BCa interval, the acceleration value \(a\) is not a
function of the bootstrap distribution and must be calculated
separately, however the process is algorithmic and can be calculated
without too much work. Each of the three previous methods (percentile,
BC, and BCa) all build upon each other and have less restrictive
assumptions, however computation increases as we loosen assumptions.

In practice, we can use the jackknife procedure to estimate the
acceleration value for this method. Thus, we will run the jackknife
procedure:

@@

Then we can compute \(a\).

@@

Now, we can find the adjust percentiles and the bias-corrected with
acceleration confidence interval.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a\_newlow }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow))}\SpecialCharTok{/}
\NormalTok{                    (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow)))); }\CommentTok{\#a\_newlow}
\NormalTok{a\_newhigh }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh))}\SpecialCharTok{/}
\NormalTok{                     (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh)))); }\CommentTok{\#a\_newhigh}
\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(a\_newlow, a\_newhigh), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

@@

@@

@@

This would give us the following histogram:

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-19-1.pdf}

@@

For the simulation, we will randomly sample from distributions in which
we know the true population mean. For example, we will sampling from a
Gamma Distribution where we can find the true population mean using the
shape and rate parameters of the sample distribution. After we sample
from a distribution, we will create a bootstrapped confidence interval
for a few of the methods we described in the exposition. Since we know
the true population mean of the distribution, we will determine if the
true population parameter is contained in the confidence interval for
each of the methods we are testing. We will perform this process a large
number of times (10,000 simulations) and then average by the number of
iterations to determines on average, the proportion that each
bootstrapped confidence interval method contains the true population
mean for a given distribution. We will do this for a couple different
distributions as well as a range of sample sizes show how the intervals
perform with skewed distributions and larger sample sizes.

@@

First, we will create a simulation that pick a random sample from a
distribution and performs

and tests to see if the sample creates a bootstrapped confidence
intervals that contain the true population parameter. First, we have the
simulation of selecting from a Normal Distribution. Since we have a
\(N(0,1)\) distribution, out true population mean is 0.

Second, we select from a Gamma Distribution with a shape parameter of 1
and a scale parameter of 4. This means the true population parameter is
4.

After that, these functions will compile all of the information we have
created from the previous methods and run the simulations of each method
for a variety of sample sizes.

And here, we run the simulations for each of the distributions and store
the results.

Then, we can plot the results and compile the results into tables.

When we sample from the Gamma distribution which is a distribution that
is heavily skewed, the Studentized Confidence interval performs the
best. At a large sample size, the Studentized Confidence interval
contains the true population parameter about 94.5\% of the time. The
Studentized confidence interval outperforms the rest of the intervals by
a bit. The rest of the intervals are very similar to each other ranging
from a correct interval 92.5\% to 93\% of the time. In general, as the
sample size that we are bootstrapping from increases, the higher
percentage our confidence intervals predict correctly.

>>>>

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge '\,'
Profossor Wagaman} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

@@

\bigskip
\begin{abstract}
This paper explores different bootstrap methods for creating confidence
intervals and tests their performance against each other for different
types of distributions. This paper will re-introduce the ideas of
bootstraps and confidence intervals, as well as the theory behind
different methods of bootstrapped confidence intervals. Along with the
theoretical details of each method, we will provide an example of each
bootstrapped confidence interval method with a toy dataset. Then, we
will run a simulation that evaluates the performance of a couple of the
methods we introduced on a variety of sample sizes and types of
distributions.
\end{abstract}

@@

Bootstrapping is an essential tool in Statistics, even more so now that
there is access to higher computing power. We use bootstrapping to
estimate the desired population parameter from a given sample without
making assumptions about any underlying distributions of the sample.
Different bootstrap techniques are developed, but we will focus on the
non-parametric bootstrap for our purposes. One way to make a statistical
inference is through confidence intervals which give a range of
estimates for the unknown population parameter at a certain confidence
level. We can use bootstrapping to create accurate approximate
confidence intervals for our population parameter even without knowing
its underlying distributions. There are many different ways to create
intervals, and we will go through a couple of them: The Standard Method,
The Percentile Method, The Bias-Corrected (BC) Method, The
Bias-Corrected with Acceleration (BCa) Method, and finally, The
Studentized Method. Additionally, we will go through an example of
creating a bootstrapped confidence interval for each method. Then, we
will perform a simulation to evaluate the performance of some of the
bootstrapped confidence interval methods on different types of
distributions.

@@

First, we introduce the non-parametric bootstrap. Suppose we have a
random sample \(X = (x_1,x_2,\dots,x_n)\) from our unknown distribution,
\(F\) and a statistic of interest, \(\hat{\theta} = \hat{\theta}(X)\)
\citep{EfronCasi}. Ideally, the desired test statistic could be found by
repeatedly sampling new reproductions of \(X\) from \(F\). However \(F\)
is unknown, so this is not possible. The non-parametric bootstrap
creates an estimate \(\hat{F}\) from \(F\) using our sample, \(X\),
without making any parametric assumptions about \(F\) (such as its
distribution type). Therefore, the bootstrap sample could be represented
as \(X^* = (x^*_1, x^*_2, \dots, x^*_n)\) where each \(x^*_i\) is
sampled randomly with equal probability and with replacement from our
original sample: \(\{x_1,x_2,\dots,x_n\}\). From this bootstrap sample,
a bootstrap replication of the test statistic can be computed using
\(\hat{\theta}^* = \hat{\theta}(X^*)\). A large number, \(B\), of
bootstrap samples are drawn independently and the corresponding
bootstrap replication of the test statistic is calculated.
\[\hat{\theta}^{*b} = \hat{\theta}(X^{*b}) \text{ for } b = 1,2, \dots, B.\]
The bootstrap estimate of the test statistic is the empirical value of
the test statistic from all of the \(\hat{\theta}^{*b}\) replications.
As \(B\) increases, \(\hat{F}\) approaches \(F\) which means that the
test statistic of interest approaches its true value as well.

\hypertarget{using-the-non-parametric-bootstrap}{%
\subsubsection{Using the Non-Parametric
Bootstrap}\label{using-the-non-parametric-bootstrap}}

We demonstrate performing a non-parametric bootstrap below using
\texttt{SnowGR} which gives the official snowfall dataset by month for
Grand Rapids, MI, starting in 1893. We will be using the \texttt{Dec}
variable which is the number of inches of snow that fell in December of
each year. In particular, we are interested in the sample mean and later
finding confidence intervals for population mean of \texttt{Dec}. The
distribution is below (Figure 1).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{data}\NormalTok{(SnowGR)}
\FunctionTok{gf\_histogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ Dec, }\AttributeTok{data =}\NormalTok{ SnowGR) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Annual Snowfall in December"}\NormalTok{, }
       \AttributeTok{title =} \StringTok{"Distribution of Annual Snowfall in December"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 1: Histogram of SnowGR"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/SnowGR histogram-1.pdf} Next,
we will perform the non-parametric bootstrap using 10,000 replications.

@@

Below, in Figure 2, is the histogram of bootstrapped values. Since we
performed a large number of bootstrap replications, the bootstrapped
distribution appears to be approximately Normal.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gf\_dhistogram}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\AttributeTok{data =}\NormalTok{ dec\_means) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{gf\_dens}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{gf\_labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Histogram of 10,000 Bootstrapped Mean Dec Values"}\NormalTok{,}
          \AttributeTok{caption =} \StringTok{"Fig. 2: Bootstrap histogram of Dec values"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. As such, the
standard confidence interval is sometimes called the normal confidence
interval for a bootstrapped sample. This means that for an unknown
skewed distribution, the standard confidence interval could present an
incorrect range. However, the same process can be used with bootstrap
sampling to form the bootstrap percentile method. This means that an
approximate bootstrap confidence interval will be created in the same
automatic way that the standard confidence interval was created. For
bootstrapped confidence intervals, the number of bootstrap replications
\(B\) must be large (around 2,000) due to the nature of confidence
intervals requiring greater accuracy \citep{Efron86}.

\hypertarget{standard-confidence-interval-in-use}{%
\subsubsection{Standard Confidence Interval in
Use}\label{standard-confidence-interval-in-use}}

Using our previous example to find a 95\% confidence interval for the
population mean of \texttt{Dec}, we would get the following confidence
interval:

@@

This means that we can say we are 95\% confident that the true mean
number of inches of snow that fall in December is betwwen 13.86 inches
and 17.66 inches. Our original histogram with the confidence interval is
below (Figure 3).

@@

@@

The percentile method interval is defined as the interval between the
\(100 * \alpha\) and the \(100(1 - \alpha)\) percentiles of the
bootstrap distribution of \(\hat{\theta}\). That is, the
\((1 - 2\alpha)\) coverage interval can be defined as
\([\hat{\theta}^*_\alpha,\hat{\theta}^*_{1-\alpha}]\)
\citep[\citet{EfronCasi}]{Efron86}. To go further, we can define
\(\hat{G}(t)\) as the bootstrap cdf, or the proportion of bootstrap
samples less than \(t\):
\[\hat{G}(t) = \frac{\#\{\hat{\theta}^{*b} \leq t\}}{B}.\] Thus the
\(\alpha\)th percentile point of the distribution is given by
\[\hat{\theta}_p[\alpha] = \hat{\theta^*_\alpha} = \hat{G}^{-1}(\alpha).\]
It follows that the percentile interval can be represented as
\[\left [ \hat{G}^{-1}(\alpha),\hat{G}^{-1}(1-\alpha) \right ].\] In the
case that the bootstrap distribution of
\(\hat{\theta}^* \sim N(\hat{\theta}, \hat{\sigma}^2)\), the
corresponding percentile interval would be equivalent to the standard
interval. However, this is not usually the case. When the bootstrap
distribution is non-normal, we can suppose that there exists, for all
\(\theta\), \[\hat{\phi} \sim N(\phi, \tau^2),\] for some monotone
transformation \(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\), and
\(\tau\) is a constant. In other words, this transformation perfectly
normalizes the distribution of \(\hat{\theta}\). This transformation
invariant can be applied to the bootstrap replications such that
\[\hat{\phi}^{*b} = g\left( \hat{\theta}^{*b}\right ) \text{ for } b = 1,2,\dots, B.\]
The corresponding percentiles of the distribution transform similarly,
\(\hat{\phi}^*_\alpha = g \left ( \hat{\theta}^*_\alpha \right )\). Or
we can say that the \((1 - 2\alpha)\) percentile interval is
\(\hat{\phi} \pm \tau z_\alpha\) which can also be represented as
\([\hat{\phi}^*_\alpha,\hat{\phi}^*_{1-\alpha}]\). This means that the
interval on the \(\theta\) scale can be defined as
\[\hat{\theta}^*_\alpha = g^{-1}(\hat{\phi} \pm \tau z_\alpha).\] This
also can be represented as an interval,
\[\left [ g^{-1}(\hat{\phi} \pm \tau z_{1-\alpha}), g^{-1}(\hat{\phi} \pm \tau z_\alpha) \right ].\]
Therefore, the percentile method produces a correct interval for
\(\phi\) and due to the transformation invariance, also produces a
correct percentile interval for \(\theta\). This method assumes the
existence of some monotone normalizing mapping
\(\hat{\phi} = g(\hat{\theta}), \phi = g(\theta)\) and relies on that to
create a correct interval. Since the process is automatic, we do not
need to know the transformation itself, only that it exists. However, in
some cases, no monotone normalizing mapping will exist \citep{Efron86}.

\hypertarget{the-percentile-method-in-use}{%
\subsubsection{The Percentile Method in
Use}\label{the-percentile-method-in-use}}

@@

Therefore, we can say that we are 95\% confident that the true
population mean of inches of snow that fall in December is between 13.91
inches and 17.72 inches. Our original histogram with the percentile
confidence interval is shown below (Figure 4).

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf} This
interval is similar to the one created using the standard method, but
shifted slightly to the right.

@@

The next method we will be looking at is the bias-corrected percentile
method (BC method) which is an improvement upon the previous percentile
method because we now take into account the possibility of bias. It can
be shown that \(\hat{\theta}\) is biased upwards relative to \(\theta\)
which means that the confidence intervals should be adjusted downwards
\citep[\citet{EfronCasi}]{Efron86}. From our simulated bootstrap
replications
\(\hat{\theta}^{*1}, \hat{\theta}^{*2}, \dots ,\hat{\theta}^{*B},\)
define \[p_0 = \frac{\#\{\hat{\theta}^{*b} \leq \theta\}}{B},\] and
define the bias-correction value \[z_0 = \Phi^{-1}(p_0),\] where
\(\Phi^{-1}\) is the inverse function of the standard normal cdf. Thus,
we define a transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\), \[\hat{\phi} \sim N(\phi - z_0\tau, \tau^2),\] with \(z_0\)
and \(\tau\) constants. This means that we can say the bias-corrected
method has an \(\alpha\)-level endpoint which can be represented as
\[\hat{\theta}_{BC}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( 2z_0 + z_\alpha\right ) \right ].\]
If \(\hat{G} = 0.50\), then half of the bootstrap distribution is less
than \(\hat{\theta}\) and our bias-correction value \(z_0 = 0\). In this
case, the confidence interval produced by BC would be the same interval
that was produced by the percentile method.

\hypertarget{the-bc-method-in-use}{%
\subsubsection{The BC Method in Use}\label{the-bc-method-in-use}}

Using this method to create a 95\% confidence interval has a couple more
steps because we need to find the bias-correction value. First we
calculate the sample mean:

@@

Then we can find the modified percentiles and get the corresponding
confidence interval:

@@

Therefore, we are 95\% confident that the true mean inches of snow that
fall in December in Grand Rapids, Michigan is between 14.01 inches and
17.82 inches. Our histogram (Figure 5) is shown below.

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

A further modification upon the BC interval is the bias-corrected and
accelerated method (BCa). For this method, we do not assume the the
standard error, \(\tau\), is constant as we did in the BC interval
\citep[\citet{EfronCasi}]{Efron86}. Rather, we assume the existence of a
monotone transformation
\(\hat{\phi} = g(\hat{\theta)}, \phi = g(\theta)\) such that for any
\(\theta\),
\[\hat{\phi} \sim N(\phi - z_0\tau_\phi, \tau_\phi^2) \text{ where } \tau_\phi = 1+ a\phi.\]
The \(a\) is known as the acceleration and is a constant that describes
how the standard deviation of \(\hat{\phi}\) varies with \(\phi\). In
other words, \(a\) is proportional to the skewness of the bootstrap
distribution. For example, \(a=z_0\) for one-parameter exponential
families, however, there are many different algorithms to compute and
estimate \(a\) \citep{Flowers18}. Now, our \(\alpha\)-level endpoint for
a BCa confidence interval is
\[\hat{\theta}_{BCa}[\alpha] = \hat{G}^{-1} \left [ \Phi \left ( z_0 + \frac{z_0+z_\alpha}{1-a(z_0+z_a)} \right ) \right ].\]
We can observe that if \(a = 0\), then
\(\hat{\theta}_{BCa}[\alpha] = \hat{\theta}_{BC}[\alpha].\) When
calculating a BCa confidence interval, the acceleration value, \(a\), is
not a function of the bootstrap distribution and must be calculated
separately. However the process is algorithmic and can be calculated
without too much work. As we saw, each of the three previous methods
(percentile, BC, and BCa) all build upon each other and have less
restrictive assumptions, but this means that computation increases as we
loosen assumptions.

\hypertarget{the-bca-confidence-interval-in-use}{%
\subsubsection{The BCa Confidence Interval in
Use}\label{the-bca-confidence-interval-in-use}}

In practice, we can use the jackknife procedure to estimate the
acceleration value for this method. The jackknife procedure is a
different resampling technique that is often used to estimate the bias
and variance of a large population. Thus, we will run the jackknife
procedure:

@@

Then we can compute an approximation for \(a\).

@@

Now, we can find the adjusted percentiles and the bias-corrected with
acceleration confidence interval.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a\_newlow }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow))}\SpecialCharTok{/}
\NormalTok{                    (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphalow)))); }
\NormalTok{a\_newhigh }\OtherTok{\textless{}{-}} \FunctionTok{pnorm}\NormalTok{(z0 }\SpecialCharTok{+}\NormalTok{ (z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh))}\SpecialCharTok{/}
\NormalTok{                     (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{estimated\_a}\SpecialCharTok{*}\NormalTok{(z0 }\SpecialCharTok{+} \FunctionTok{qnorm}\NormalTok{(alphahigh)))); }
\FunctionTok{qdata}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{ mean, }\FunctionTok{c}\NormalTok{(a\_newlow, a\_newhigh), }\AttributeTok{data =}\NormalTok{ dec\_means)}
\end{Highlighting}
\end{Shaded}

@@

We are 95\% confident that the true mean number of inches of snow that
fall in December is between 13.92 inches and 17.72 inches. Figure 6
shows our confidence interval.

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

So we are 95\% confident that the true mean number of inches of snow
that fall in December in Grand Rapids, Michigan is between 14.03 inches
and 18.02 inches. This would give us the following confidence interval
(Figure 7)

@@

@@

For our simulation, we will randomly sample from distributions in which
we know the true population mean. For example, we will be sampling from
a Gamma distribution where we can find the true population mean using
the shape and rate parameters of the sample distribution. After we
sample from the distribution, we will create a bootstrapped confidence
interval for a few of the methods we described in the exposition. Since
we know the true population mean of the distribution, we will determine
if the true population mean is contained in the confidence interval for
each of the methods we are testing. We will perform this process a large
number of times (10,000 simulations) and then average by the number of
iterations to determines on average, the proportion that each
bootstrapped confidence interval method contains the true population
mean for a given distribution. We will do this for a couple different
distributions as well as a range of sample sizes to show how the
intervals perform with skewed distributions and larger sample sizes.

@@

To start, we will declare some constants including the sample sizes and
the number of bootstrap replications we are using.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_sizes }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{60}\NormalTok{, }\DecValTok{100}\NormalTok{, }\DecValTok{150}\NormalTok{)}
\NormalTok{num\_sample\_sizes }\OtherTok{\textless{}{-}} \FunctionTok{length}\NormalTok{(sample\_sizes)}
\NormalTok{boot\_reps }\OtherTok{\textless{}{-}} \DecValTok{2000}
\end{Highlighting}
\end{Shaded}

Next, we create a simulation that randomly selects a sample from a
Normal distribution and performs a bootstrap with 2,000 bootstrap
replications. Then, we create the confidence intervals from this
bootstrap and determine if the true mean, our population parameter, is
contained within our intervals. Since we are using a \(Normal(0,1)\)
distribution, our true population mean is equal to 0. We will repeat
this process 10,000 times as specified for each of the sample sizes that
we are testing. Since we are originally sampling from a Normal
distribution, the assumptions we hold for the standard interval are true
and thus, we would expect all of the methods to perform comparably to
each other at large sample sizes.
